Variational Monte Carlo on bosonic systems

### 2018-02-07 project_1/article/article.tex
Add derivation of the interaction terms
	Could you try to reproduce equations 40 and 41 @gregwinther?

                we can state an intermediate version of the local energy as

@@

            component for particle $k$ by

@@

                \tpl{{r_k}}
                = \dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}.

@@

            The derivative of the distance $r_{km}$ is given by
            \begin{gather}
                r_{km} = |\vf{r}_k - \vf{r}_m|
                \implies
                \dpd[]{r_{km}}{r_k}
                = \frac{r_k - r_m}{|\vf{r}_k - \vf{r}_m|},
            \end{gather}
            where in spherical coordinates $\vf{r}_k = r_k\vf{e}_r$.  The
            gradient of $u(r_{km})$ thus simplifies to only include the radial
            contribution.
                \nabla_k u(r_{km})
                \vf{e}_{r}\tpl{{r_k}}u(r_{km})
                \vf{e}_r\dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}u(r_{km})
                &= \vf{e}_r\frac{r_k - r_m}{r_{km}}
                \tpl{{r_{km}}}u(r_{km}).
                Using this expression we find the Laplacian to be
                    \nabla_k^2 u(r_{km})
                    \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
                    + \frac{(r_k - r_m)^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
                    \frac{r_k - r_m}{r_{km}}
                    \dpd[]{r_{km}}{r_k}
                    \partial_{{r_{km}}}^2 u(r_{km}) \\
                    \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
                    + \frac{r_{km}^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
                    \frac{(r_k - r_m)^2}{r_{km}^2}
                    \partial_{{r_{km}}}^2 u(r_{km}) \\
                    \frac{2}{r_{km}}\tpl{{r_{km}}}u(r_{km})
                    + \partial_{{r_{km}}}^2u(r_{km}).
            We now take the derivate with respect to $r_{km}$ of the interaction
            functions $u(r_{km})$ to find the closed form expressions. We get
                \partial_{r_{km}}u(r_{km})
                \partial_{r_{km}}
                \ln\Biggl(
                    1 - \frac{a}{r_{km}}
                \Biggr) \\
                \partial_{r_{km}}
                \ln\Biggl(
                    \frac{1}{r_{km}}
                    (r_{km} - a)
                \Biggr) \\
                \partial_{r_{km}}
                    \ln(r_{km} - a)
                    - \ln(r_{km})
                \Bigr] \\
                \frac{1}{r_{km} - a} - \frac{1}{r_{km}} \\
                \frac{a}{r_{km}(r_{km} - a)} \\
                \partial_{r_{km}}^2u(r_{km})
                &= a\Bigl(
                    \partial_{r_{km}}\frac{1}{r_{km}}
                \Bigr)
                \partial_{r_{km}}\frac{1}{r_{km} - a}
                + a\frac{1}{r_{km}}
                \Bigl(
                    \partial_{r_{km}}\frac{1}{r_{km} - a}
                \Bigr) \\
                -\frac{a}{r_{km}^2(r_{km} - a)}
                -\frac{a}{r_{km}(r_{km} - a)^2} \\
                - \frac{2a(r_{km} - 1)}{r_{km}^2(r_{km} - a)^2}.
            Using this we are now able to write the closed form expression of
            the Laplacian of the trial wavefunction.
                    \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
                    {\bra{\vf{r}}\ket{\Psi_T}}
                    2a
                    \Biggl[
                        \frac{1}{r_{km}^2(r_{km} - a)}
                        - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
                    -4\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
                    \cdot
                    a\sum_{m \neq k}^N
                    \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
                        a\sum_{m \neq k}^N
                        \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
                    \Biggr]^2.
            To ease the load on the CPU we try to restrict the number of times
            we evaluate the sums. For brevity we introduce the following
            functions
                \xi_k(\vf{r}, a)
                &= \sum_{m \neq k}^N
                \Biggl[
                    \frac{1}{r_{km}^2(r_{km} - a)}
                    - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
                \Biggr] \\
                \frac{1 - a}{r_{km}^2(r_{km} - a)^2}, \\
                \vfg{\zeta}_k(\vf{r}, a)
                \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}.
            We can then rewrite the expression for the Laplacian to
                    \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
                    {\bra{\vf{r}}\ket{\Psi_T}}
                    2a\xi_k(\vf{r}, a)
                    -4a\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
                    \cdot
                    \vfg{\zeta}_k(\vf{r}, a)
                    + a^2\vfg{\zeta}_k(\vf{r}, a)^2.
                Both $\xi_k$ and $\vfg{\zeta}_k$ can be evaluated in a common
                loop. Using this expression for the Laplacian we can find the
                local energy from
                    E_L(\vf{r})
                    \frac{\bra{\vf{r}}H\ket{\Psi_T}}
                    {\bra{\vf{r}}\ket{\Psi_T}}
                    \sum_{k}^N
                        -\frac{\hbar^2}{2m}
                        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
                        {\bra{\vf{r}}\ket{\Psi_T}}
                        + V_{\text{ext}}(\vf{r})
                    \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j).

>>>>

                we can state an intermediate version of the Laplacian occuring
                in the local energy as

@@

            Moving on to the derivatives of the interaction terms, $u_{km}$, to
            get an explicit expression for the Laplacian.

@@

                \frac{a}{r_{km}(r_{km} - a)},
                \dpd[2]{u_{km}}{r_{km}}
                &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.

@@

            %component for particle $k$ by
            %\begin{align}
            %    \tpl{{r_k}}
            %    = \dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}.
            %\end{align}
            %The derivative of the distance $r_{km}$ is given by
            %\begin{gather}
            %    r_{km} = |\vf{r}_k - \vf{r}_m|
            %    \implies
            %    \dpd[]{r_{km}}{r_k}
            %    = \frac{r_k - r_m}{|\vf{r}_k - \vf{r}_m|},
            %\end{gather}
            %where in spherical coordinates $\vf{r}_k = r_k\vf{e}_r$.  The
            %gradient of $u(r_{km})$ thus simplifies to only include the radial
            %contribution.
            %\begin{align}
            %    \nabla_k u(r_{km})
            %    \vf{e}_{r}\tpl{{r_k}}u(r_{km})
            %    \vf{e}_r\dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}u(r_{km})
            %    &= \vf{e}_r\frac{r_k - r_m}{r_{km}}
            %    \tpl{{r_{km}}}u(r_{km}).
            %\end{align}
            %    Using this expression we find the Laplacian to be
            %        \nabla_k^2 u(r_{km})
            %        \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
            %        + \frac{(r_k - r_m)^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
            %        \frac{r_k - r_m}{r_{km}}
            %        \dpd[]{r_{km}}{r_k}
            %        \partial_{{r_{km}}}^2 u(r_{km}) \\
            %        \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
            %        + \frac{r_{km}^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
            %        \frac{(r_k - r_m)^2}{r_{km}^2}
            %        \partial_{{r_{km}}}^2 u(r_{km}) \\
            %        \frac{2}{r_{km}}\tpl{{r_{km}}}u(r_{km})
            %        + \partial_{{r_{km}}}^2u(r_{km}).
            %We now take the derivate with respect to $r_{km}$ of the interaction
            %functions $u(r_{km})$ to find the closed form expressions. We get
            %\begin{align}
            %    \partial_{r_{km}}u(r_{km})
            %    \partial_{r_{km}}
            %    \ln\Biggl(
            %        1 - \frac{a}{r_{km}}
            %    \Biggr) \\
            %    \partial_{r_{km}}
            %    \ln\Biggl(
            %        \frac{1}{r_{km}}
            %        (r_{km} - a)
            %    \Biggr) \\
            %    \partial_{r_{km}}
            %    \Bigl[
            %        \ln(r_{km} - a)
            %        - \ln(r_{km})
            %    \Bigr] \\
            %    \frac{1}{r_{km} - a} - \frac{1}{r_{km}} \\
            %    \frac{a}{r_{km}(r_{km} - a)} \\
            %    \partial_{r_{km}}^2u(r_{km})
            %    &= a\Bigl(
            %        \partial_{r_{km}}\frac{1}{r_{km}}
            %    \Bigr)
            %    \partial_{r_{km}}\frac{1}{r_{km} - a}
            %    \nonumber \\
            %    + a\frac{1}{r_{km}}
            %    \Bigl(
            %        \partial_{r_{km}}\frac{1}{r_{km} - a}
            %    \Bigr) \\
            %    -\frac{a}{r_{km}^2(r_{km} - a)}
            %    -\frac{a}{r_{km}(r_{km} - a)^2} \\
            %    - \frac{2a(r_{km} - 1)}{r_{km}^2(r_{km} - a)^2}.
            %\end{align}
            %Using this we are now able to write the closed form expression of
            %the Laplacian of the trial wavefunction.
            %        -2\alpha
            %        \big(
            %            2 + \beta
            %        \bigr)
            %        + 4\alpha^2
            %        \bigl(
            %            x_k^2 + y_k^2 + \beta^2z_k^2
            %        \bigr)
            %        %\nonumber \\
            %        %&\qquad
            %        2a
            %        \sum_{m \neq k}^N
            %        \Biggl[
            %            \frac{1}{r_{km}^2(r_{km} - a)}
            %            - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
            %        \Biggr]
            %        \nonumber \\
            %        -4\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
            %        \cdot
            %        a\sum_{m \neq k}^N
            %        \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
            %        + \Biggl[
            %            a\sum_{m \neq k}^N
            %            \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
            %        \Biggr]^2.
            %To ease the load on the CPU we try to restrict the number of times
            %we evaluate the sums. For brevity we introduce the following
            %functions
            %\begin{align}
            %    \xi_k(\vf{r}, a)
            %    &= \sum_{m \neq k}^N
            %    \Biggl[
            %        \frac{1}{r_{km}^2(r_{km} - a)}
            %        - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
            %    \Biggr] \\
            %    \sum_{m \neq k}^N
            %    \frac{1 - a}{r_{km}^2(r_{km} - a)^2}, \\
            %    \vfg{\zeta}_k(\vf{r}, a)
            %    \sum_{m \neq k}^N
            %    \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}.
            %\end{align}
            %We can then rewrite the expression for the Laplacian to
            %        -2\alpha
            %        \big(
            %            2 + \beta
            %        \bigr)
            %        + 4\alpha^2
            %        \bigl(
            %            x_k^2 + y_k^2 + \beta^2z_k^2
            %        \bigr)
            %        %\nonumber \\
            %        %&\qquad
            %        2a\xi_k(\vf{r}, a)
            %        \nonumber \\
            %        -4a\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
            %        \cdot
            %        \vfg{\zeta}_k(\vf{r}, a)
            %        + a^2\vfg{\zeta}_k(\vf{r}, a)^2.
            %    Both $\xi_k$ and $\vfg{\zeta}_k$ can be evaluated in a common
            %    loop. Using this expression for the Laplacian we can find the
            %    local energy from
            %        E_L(\vf{r})
            %        \frac{\bra{\vf{r}}H\ket{\Psi_T}}
            %        =
            %        \sum_{k}^N
            %            -\frac{\hbar^2}{2m}
            %            + V_{\text{ext}}(\vf{r})
            %        \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j).
